---
title: "Cox Model"
author: Divya Veerapaneni MS4, Ong Lab
input: preprocessed cox dataset (output from Preprocess_Cox_Dataset.ipynb)
output: html_notebook
---
### Predicts Survival Probability from hospitalization in Traumatic Brain Injury Patients using Pupil Metrics from Multiple Observations for each patients using a Cox Model
#### Pupil Metrics of Interest = Average NPi, Min NPi, NPi Difference, Size Difference

```{r}
#import statements 
library(readxl)
library(geepack)
library(broom)
library(tidyverse)
library(lubridate)
library(readxl)
library(survival)
library(dplyr)
library(ggplot2)
library(gridExtra)
library("survminer")
options(scipen=999) #prevents outputs in scientific notation 

#input dataframe
df <- read.csv("Cox_model.csv")

factor_variables = list('RACE_Black', 'tbi_severity', 'Mechanism_Injury','ID', 'any_incidence', 'poor_npi_incidence', 'npi_diff_incidence',
       'size_diff_incidence', 'uni_any_incidence', 'stage1u_incidence',
       'stage2u_incidence', 'stage3u_incidence', 'stage4u_incidence',
       'bi_incidence', 'stage1b_incidence', 'stage2b_incidence',
       'stage3b_incidence', 'npi_diff_size_diff_incidence',
       'size_diff_poor_npi_incidence')

#factor each variable in factor_variables
for (variable in factor_variables)
{
   df[, c(variable)] <- as.factor(df[, c(variable)])
}
df[['date']] <- as.POSIXct(df[['date']],format = "%Y-%m-%d %H:%M:%S")
```

# Survival Plots 

### Deceased vs Average NPi
```{r}
# convert average NPi to dichotomous categorical variable (NPi 0-3 vs NPi 3-5)
df$average_npi_cat <- cut(df$average_npi, breaks=c(0, 3, 5), labels=c("NPi 0-3", "NPi 3-5"), include.lowest = TRUE)

# create survival object
fit <- survfit(Surv(time_start, time_end, Deceased_cox) ~ average_npi_cat, data = df)

# plot survival plot (error when trying to plot p value on right-censored data)
ggsurvplot(fit, data=df,title="Kaplan-Meier Curve for Survival in TBI Patients by Average NPi",
           legend.labs=c("[0-3]", "(3-5]"), legend.title="Average NPi:")
```
# Univariate Analysis
```{r}
explanatory_vars =  c( 'average_npi', 'lower_npi', 'average_npi_0_removed',
       'lower_npi_0_removed', 'npi_diff', 'size_diff',   'any_incidence', 'poor_npi_incidence', 'npi_diff_incidence',
       'size_diff_incidence', 'uni_any_incidence', 'stage1u_incidence',
       'stage2u_incidence', 'stage3u_incidence', 'stage4u_incidence',
       'bi_incidence', 'stage1b_incidence', 'stage2b_incidence',
       'stage3b_incidence', 'npi_diff_size_diff_incidence',
       'size_diff_poor_npi_incidence')

cox_model_list <- map(explanatory_vars, 
                   ~coxph(reformulate(.x, "Surv(time_start, time_end, Deceased_cox)"), data = df))

uni_tb <- map_df(cox_model_list, broom::tidy, exponentiate = TRUE, conf.int=TRUE) %>%
mutate(HR = round(estimate,3), ci_low = round(exp(conf.low),2), ci_high = round(exp(conf.high),2), p_value = round(p.value,3), CI = paste(ci_low, ci_high, sep = "-")) %>%
select(term, HR, CI, p_value)
uni_tb
```

# Multivariable Analysis
### Deceased vs average_NPi adjusted for age, sex, race, Marshall score, and TBI severity
```{r}
#cox model
cox_model_list <- map(explanatory_vars, 
                   ~coxph(formula(paste("Surv(time_start, time_end, Deceased_cox) ~", .x,  " + AGE +RACE_Black+tbi_severity + Mechanism_Injury + Marshall")), data = df))

mv_tb <- map_df(cox_model_list, broom::tidy, exponentiate = TRUE, conf.int=TRUE) %>%
mutate(HR = round(estimate,3), ci_low = round(exp(conf.low),2), ci_high = round(exp(conf.high),2), p_value = round(p.value,3), CI = paste(ci_low, ci_high, sep = "-")) %>%
filter(row_number()%%7 == 1) %>%
select(term, HR, CI, p_value)
mv_tb
```


