---
title: "Multivariable Analysis of Pupil Abnormalities and Discharge Outcomes"
author: Divya Veerapaneni MS4, Ong Lab
input: Consolidated_Study_Cohort_Traits.xlsx - excel sheet with final patient data including demographic data, pupil metrics, incidence and frequency of abnormal pupil phenotypes 
output:
  html_document:
    df_print: paged
---

# Multivariable Analysis of Pupil Abnormalities and Discharge Outcomes
### y = Unfavorable Outcomes, Deceased, and Awake/alert
### x = mean NPi, min NPi, NPI diff, size diff, stages incidence and burden
### confounding variables = Age, Sex, Black race, TBI severity, GCS, blunt vs penetrating injury, Marshall CT score

#### Preprocessing Data
```{r}
#import statements
library(readxl)
library(gtsummary)
library(tidyverse)
library(broom)
library(flextable)
library(gridExtra)
library(logistf)

#input file
df <- read_excel("/Users/divs/Box/1-BMC Smartguards/10-Processing and Visualization/8-TBI Pupillometry/Data/Consolidated_Study_Cohort_Traits.xlsx")

#variables of interest in input file
factor_variables = list("Unfavorable_Outcome", "Orientedx3", "Awake_Alert", "Deceased", "RACE_Black", "tbi_severity",
    "Mechanism_Injury", "Marshall",
     "stage1u_incidence", "stage2u_incidence", "stage3u_incidence", "stage1b_incidence", "stage2b_incidence", 
     "uni_incidence", "bi_incidence", "any_incidence" )
```

#### Multivariable Analysis when adjusting for Age, Black Race, TBI severity, Mechanism of Injury, and Marshall Scale On Full Cohort
```{r}

explanatory_vars <- c('average_npi', 'lower_npi', 'npi_diff', 'size_diff', 
 'any_incidence', 'poor_npi_incidence','npi_diff_incidence', 'size_diff_incidence',
 'uni_any_incidence', 'stage1u_incidence', 'stage2u_incidence', 'stage3u_incidence',
       'bi_incidence', 'stage1b_incidence','stage2b_incidence', 'stage3b_incidence', 
  'any_burden', 'poor_npi_burden', 'npi_diff_burden', 'size_diff_burden',
 'uni_any_burden','stage1u_burden', 'stage2u_burden', 'stage3u_burden', 
        'bi_burden', 'stage1b_burden', 'stage2b_burden','stage3b_burden',
 'average_npi_0_removed', 'lower_npi_0_removed', 
 'stage4u_incidence', 'npi_diff_size_diff_incidence', 'size_diff_poor_npi_incidence', 
 'stage4u_burden',  'npi_diff_size_diff_burden','size_diff_poor_npi_burden')

create_multivariable_table <- function(data_df, y, x_list){
  tb <- 
map_df(explanatory_vars,
       ~glm(formula(paste(y, "~", .x,  " + AGE +RACE_Black + tbi_severity + Mechanism_Injury + Marshall")), data = data_df, family = binomial) %>%
         tidy(conf.int = TRUE)) %>%
  filter(term %in% explanatory_vars) %>%
  mutate(OR = round(exp(estimate),2), ci_low = round(exp(conf.low),2), ci_high = round(exp(conf.high),2), p_value = round(p.value,3), CI = paste(ci_low, ci_high, sep = "-")) %>%
  select(term, OR, CI, p_value) %>%
  rename(!! paste0(y) := term, OR=OR, CI=CI, p_value=p_value)
  return (tb)
}
UO_tb <- create_multivariable_table(df, "Unfavorable_Outcome", explanatory_vars)  
Deceased_tb <- create_multivariable_table(df, "Deceased", explanatory_vars)
Awake_tb <- create_multivariable_table(df, "Awake_Alert", explanatory_vars)
#Oriented_tb <- create_multivariable_table(df, "Orientedx3", explanatory_vars)
cbind(UO_tb, Deceased_tb, Awake_tb)

#write.csv(dat_append, "/Users/divs/Downloads/Multivariate_Marshall_11-15.csv")

#png("/Users/divs/Downloads/Multivariate_full_cohort.png", height = 25*nrow(dat_append), width = 150*ncol(dat_append))
#grid.table(dat_append)
#dev.off()

```
# Multivariable Analysis when adjusting for Age, Black Race, TBI severity, Mechanism of Injury, and Marshall Scale On Abnormal Patient Cohort (n=102)
```{r}

df_n102 <-  filter(df, any_incidence == 1)
create_multivariable_table <- function(data_df, y, x_list){
  tb <- 
map_df(explanatory_vars,
       ~glm(formula(paste(y, "~", .x,  " + AGE +RACE_Black+tbi_severity + Mechanism_Injury + Marshall")), data = data_df, family = binomial) %>%
         tidy(conf.int = TRUE)) %>%
  filter(term %in% explanatory_vars) %>%
  mutate(OR = round(exp(estimate),2), ci_low = round(exp(conf.low),2), ci_high = round(exp(conf.high),2), p_value = round(p.value,3), CI = paste(ci_low, ci_high, sep = "-")) %>%
  select(term, OR, CI, p_value) %>%
  rename(!! paste0(y) := term, OR=OR, CI=CI, p_value=p_value)
  return (tb)
}

UO_tb <- create_multivariable_table(df_n102, "Unfavorable_Outcome", explanatory_vars)  
Deceased_tb <- create_multivariable_table(df_n102, "Deceased", explanatory_vars)
Awake_tb <- create_multivariable_table(df_n102, "Awake_Alert", explanatory_vars)
#Oriented_tb <- create_multivariable_table(df_n102, "Orientedx3", explanatory_vars)
cbind(UO_tb, Deceased_tb, Awake_tb)

#write.csv(dat_append_n102, "/Users/divs/Downloads/Multivariate_102_Marshall_11-15.csv")

#png("/Users/divs/Downloads/Multivariate_full_cohort.png", height = 25*nrow(dat_append_n102), width = 150*ncol(dat_append_n102))
#grid.table(dat_append_n102)
#dev.off()

```

# Multivariable Firth For Incidence Where Less Than 10 Patients
```{r}
vars <-   list('stage2b_incidence', 'stage3b_incidence')

outcomes <- list('Unfavorable_Outcome', 'Deceased', 'Awake_Alert')
firth_final = data.frame()
for (v in vars){
  print(v)
  row_final = data.frame(v) # create dataframe for row
  for (outcome in outcomes){
   
    mod <- logistf(formula(paste(outcome, "~", v, "+ AGE +RACE_Black + tbi_severity + Mechanism_Injury + Marshall")), data = df) #create model  
    OR <- round(exp(mod$coefficients[2]),2) # get OR for 2nd term rounded to 2 decimal points
    CI <- round(exp(confint(mod, paste0(v)))[2,],2) # get confidence interval for 2nd term
    CI <- paste(CI[1], CI[2], sep='-') # format CI to ci.low-ci-high
    p <- round(mod$prob[2],3) # get p value
    temp <- data.frame(outcome,OR,CI, p) #row-wise data
    row_final <- cbind(row_final,temp) # combine all outcome results for same v into single row
  }
  firth_final <- rbind(firth_final, row_final) # add row by row as v changes
}
firth_final[ -c(1) ] # remove dummy first row
# write.csv(firth_final, "/Users/divs/Downloads/Firth_Multivariable_Nonzero_Burden.csv")
```

# Multivariable Firth For All Burden Where Less Than 10 Patients
```{r}
vars <-   list('stage2b_burden','stage3b_burden')

outcomes <- list('Unfavorable_Outcome', 'Deceased', 'Awake_Alert')

v <- "stage2b_burden"
outcome <- "Unfavorable_Outcome"

if (v == 'stage2b_burden' ){#| v == 'stage3b_burden'){  
      mod <- logistf(formula(paste(outcome, "~", v, "+ AGE +RACE_Black + Mechanism_Injury  + Marshall")), data = df) #create modelprint(v)
    }else{
      mod <- logistf(formula(paste(outcome, "~", v, "+ AGE +RACE_Black + tbi_severity + Mechanism_Injury + Marshall")), data = df) #create model  
      print(v)
    }
    
    OR <- round(exp(mod$coefficients[2]),2) # get OR for 2nd term rounded to 2 decimal points
    CI <- round(exp(confint(mod, paste0(v)))[2,],2) # get confidence interval for 2nd term
    CI <- paste(CI[1], CI[2], sep='-') # format CI to ci.low-ci-high
    p <- round(mod$prob[2],3) # get p value
    data.frame(outcome,OR,CI, p) #row-wise data
```

# Moderate TBI Multivariable Subanalysis
```{r}

create_multivariable_table2 <- function(data_df, y, x_list, formula){
  tb <- 
map_df(explanatory_vars,
       ~glm(formula(paste(y, "~", .x,  formula)), data = data_df, family = binomial) %>%
         tidy(conf.int = TRUE)) %>%
  filter(term %in% explanatory_vars) %>%
  mutate(OR = round(exp(estimate),2), ci_low = round(exp(conf.low),2), ci_high = round(exp(conf.high),2), p_value = round(p.value,3), CI = paste(ci_low, ci_high, sep = "-")) %>%
  select(term, OR, CI, p_value) %>%
  rename(!! paste0(y) := term, OR=OR, CI=CI, p_value=p_value)
  return (tb)
}
#create results for mod TBI patients
modTBI_df <- filter(df, tbi_severity == "Moderate")

UO_tbi_mod <- create_multivariable_table2(modTBI_df, "Unfavorable_Outcome", explanatory_vars,"+ AGE + RACE_Black + Mechanism_Injury + Marshall" )  
Deceased_tbi_mod <- create_multivariable_table2(modTBI_df, "Deceased", explanatory_vars, "+ AGE + RACE_Black  + Marshall" )
Awake_tbi_mod <- create_multivariable_table2(modTBI_df, "Awake_Alert", explanatory_vars, "+ AGE + RACE_Black  + Marshall")
cbind(UO_tbi_mod, Deceased_tbi_mod, Awake_tbi_mod)
#write.csv(dat_append_mod, "/Users/divs/Downloads/Multivariable_TBI_Marshall_11-15.csv")
```

# Severe TBI Multivariable Subanalysis
```{r}
#create results for severe TBI patients
severeTBI_df <- filter(df, tbi_severity == "Severe")

UO_tbi_severe <- create_multivariable_table2(severeTBI_df, "Unfavorable_Outcome", explanatory_vars,"+ AGE + RACE_Black + Mechanism_Injury + Marshall" )  
Deceased_tbi_severe <- create_multivariable_table2(severeTBI_df, "Deceased", explanatory_vars, "+ AGE + RACE_Black + Mechanism_Injury + Marshall" )
Awake_tbi_severe <- create_multivariable_table2(severeTBI_df, "Awake_Alert", explanatory_vars, "+ AGE + RACE_Black + Mechanism_Injury + Marshall")
cbind(UO_tbi_severe, Deceased_tbi_severe, Awake_tbi_severe)
#write.csv(dat_append_severe, "/Users/divs/Downloads/Multivariable_severeTBI_Marshall_11-15.csv")
```




