---
title: "Multivariable Analysis of Pupil Abnormalities and Discharge Outcomes"
author: Divya Veerapaneni MS4, Ong Lab
input: Consolidated_Study_Cohort_Traits.xlsx - excel sheet with final patient data including demographic data, pupil metrics, incidence and frequency of abnormal pupil phenotypes 
output:
  html_document:
    df_print: paged
---

# Multivariable Analysis of Pupil Abnormalities and Discharge Outcomes
### y = Unfavorable Outcomes, Deceased, and Awake/alert
### x = mean NPi, min NPi, NPI diff, size diff, stages incidence and burden
### confounding variables = Age, Sex, Black race, TBI severity, GCS, blunt vs penetrating injury, Marshall CT score

#### Preprocessing Data
```{r}
#import statements
library(readxl)
library(gtsummary)
library(tidyverse)
library(broom)
library(flextable)
#library(gridExtra)
library(logistf)

#input file
df <- read_excel("/Users/divs/Box/1-BMC Smartguards/10-Processing and Visualization/8-TBI Pupillometry/Data/Consolidated_Study_Cohort_Traits.xlsx")

#variables of interest in input file
factor_variables = list("Unfavorable_Outcome", "Awake_Alert", "Deceased", "RACE", "tbi_severity",
    "Mechanism_Injury", "Marshall",
     "uni_incidence", "bi_incidence", "any_incidence" )
df$RACE <- as.factor(df$RACE) %>%
relevel("White")
```

```{r}
# Use for Incidence
create_multivariable_table_incidence <- function(data_df, y, x_list){
  tb <- 
map_df(explanatory_vars,
       ~glm(formula(paste(y, "~", .x,  " + AGE +RACE + tbi_severity + Mechanism_Injury + Marshall")), data = data_df, family = binomial) %>%
         tidy(conf.int = TRUE)) %>%
  filter(term %in% explanatory_vars) %>%
  mutate(OR = round(exp(estimate),2), ci_low = round(exp(conf.low),2), ci_high = round(exp(conf.high),2), p_value = round(p.value,3), CI = paste(ci_low, ci_high, sep = "-")) %>%
  select(term, OR, CI, p_value) %>%
  rename(!! paste0(y) := term, OR=OR, CI=CI, p_value=p_value)
  return (tb)
}

# Use for Burden For 1% Increase
create_multivariable_table_OR1 <- function(data_df, y, x_list){
  percent_increase <- 1
  tb <- 
map_df(explanatory_vars,
       ~glm(formula(paste(y, "~", .x,  " + AGE +RACE + tbi_severity + Mechanism_Injury + Marshall")), data = data_df, family = binomial) %>%
         tidy(conf.int = TRUE)) %>%
  filter(term %in% explanatory_vars) %>%
  mutate(OR = round(exp(estimate*percent_increase),2), ci_low = round(exp(conf.low*percent_increase),2), ci_high = round(exp(conf.high*percent_increase),2), p_value = round(p.value,3), CI = paste(ci_low, ci_high, sep = "-")) %>%
  select(term, OR, CI, p_value) %>%
  rename(!! paste0(y) := term, OR=OR, CI=CI, p_value=p_value)
  return (tb)
}

# Use for Incidence in Severe TBI
create_multivariable_table_incidence_severe <- function(data_df, y, x_list){
  tb <- 
map_df(explanatory_vars,
       ~glm(formula(paste(y, "~", .x,  " + AGE +RACE + Mechanism_Injury + Marshall")), data = data_df, family = binomial) %>%
         tidy(conf.int = TRUE)) %>%
  filter(term %in% explanatory_vars) %>%
  mutate(OR = round(exp(estimate),2), ci_low = round(exp(conf.low),2), ci_high = round(exp(conf.high),2), p_value = round(p.value,3), CI = paste(ci_low, ci_high, sep = "-")) %>%
  select(term, OR, CI, p_value) %>%
  rename(!! paste0(y) := term, OR=OR, CI=CI, p_value=p_value)
  return (tb)
}

# Use for Burden in Severe TBI
create_multivariable_table_OR1_severe <- function(data_df, y, x_list){
  percent_increase <- 1
  tb <- 
map_df(explanatory_vars,
       ~glm(formula(paste(y, "~", .x,  " + AGE +RACE + Mechanism_Injury + Marshall")), data = data_df, family = binomial) %>%
         tidy(conf.int = TRUE)) %>%
  filter(term %in% explanatory_vars) %>%
  mutate(OR = round(exp(estimate*percent_increase),2), ci_low = round(exp(conf.low*percent_increase),2), ci_high = round(exp(conf.high*percent_increase),2), p_value = round(p.value,3), CI = paste(ci_low, ci_high, sep = "-")) %>%
  select(term, OR, CI, p_value) %>%
  rename(!! paste0(y) := term, OR=OR, CI=CI, p_value=p_value)
  return (tb)
}
```


#### Multivariable Analysis when adjusting for Age, Black Race, TBI severity, Mechanism of Injury, and Marshall Scale On Full Cohort by 3% Increase 
```{r}

explanatory_vars <- c(
  'any_burden', 'poor_npi_burden', 'npi_diff_burden', 'size_diff_burden',
 'uni_any_burden','bi_burden')

UO_tb <- create_multivariable_table_OR1(df, "Unfavorable_Outcome", explanatory_vars)  
Deceased_tb <- create_multivariable_table_OR1(df, "Deceased", explanatory_vars)
Awake_tb <- create_multivariable_table_OR1(df, "Awake_Alert", explanatory_vars)
MV_Burden_OR1_Results <- cbind(UO_tb, Deceased_tb, Awake_tb)
write.csv(MV_Burden_OR1_Results, "/Users/divs/Downloads/Multivariable_OR1_freq.csv")
```

#### Multivariable Analysis when adjusting for Age, Black Race, TBI severity, Mechanism of Injury, and Marshall Scale On Full Cohort for 1% Increase
```{r}

explanatory_vars <- c('average_npi', 'lower_npi', 'npi_diff', 'size_diff', 
 'any_incidence', 'poor_npi_incidence','npi_diff_incidence', 'size_diff_incidence',
 'uni_any_incidence', 'bi_incidence')

UO_tb <- create_multivariable_table_incidence(df, "Unfavorable_Outcome", explanatory_vars)  
Deceased_tb <- create_multivariable_table_incidence(df, "Deceased", explanatory_vars)
Awake_tb <- create_multivariable_table_incidence(df, "Awake_Alert", explanatory_vars)
MV_Incidence_OR1_Results <- cbind(UO_tb, Deceased_tb, Awake_tb)
write.csv(MV_Burden_OR1_Results, "/Users/divs/Downloads/Multivariable_OR1_incidence.csv")
```



# Severe TBI Multivariable Subanalysis
```{r}
#create results for severe TBI patients
severeTBI_df <- filter(df, tbi_severity == "Severe")


#first compute analysis for incidence
explanatory_vars <- c(
  'average_npi', 'lower_npi', 'npi_diff', 'size_diff', 
  'any_incidence','uni_any_incidence', 'bi_incidence')

UO_tbi_severe <- create_multivariable_table_incidence_severe(severeTBI_df, "Unfavorable_Outcome", explanatory_vars)  
Deceased_tbi_severe <- create_multivariable_table_incidence_severe(severeTBI_df, "Deceased", explanatory_vars)
Awake_tbi_severe <- create_multivariable_table_incidence_severe(severeTBI_df, "Awake_Alert", explanatory_vars)
MV_severe_incidence <- cbind(UO_tbi_severe, Deceased_tbi_severe, Awake_tbi_severe)
write.csv(MV_severe_incidence, "/Users/divs/Downloads/MV_severe_incidence.csv")

#then compute analysis for burden
explanatory_vars <- c(
  'any_burden','uni_any_burden', 'bi_burden')
UO_tbi_severe <- create_multivariable_table_OR1_severe(severeTBI_df, "Unfavorable_Outcome", explanatory_vars)  
Deceased_tbi_severe <- create_multivariable_table_OR1_severe(severeTBI_df, "Deceased", explanatory_vars)
Awake_tbi_severe <- create_multivariable_table_OR1_severe(severeTBI_df, "Awake_Alert", explanatory_vars)
MV_severe_burden <- cbind(UO_tbi_severe, Deceased_tbi_severe, Awake_tbi_severe)
write.csv(MV_severe_burden, "/Users/divs/Downloads/Severe_OR1_Freq.csv")

```




