---
title: "Chi-Square Analysis of Categorical Variables and TBI Severity"
author: Divya Veerapaneni MS4, Ong Lab
input: Consolidated_Study_Cohort_Traits.xlsx - excel sheet with final patient data including demographic data, pupil metrics, incidence and frequency of abnormal pupil phenotypes 
output:
  html_document:
    df_print: paged
---

#### Preprocessing Data
```{r}
#import statements
library(readxl)
library(gtsummary)
library(tidyverse)
library(broom)
library(flextable)
#library(gridExtra)
library(logistf)

#input file
df <- read_excel("/Users/divs/Box/1-BMC Smartguards/10-Processing and Visualization/8-TBI Pupillometry/Data/Consolidated_Study_Cohort_Traits.xlsx")

#variables of interest in input file
factor_variables = list("Unfavorable_Outcome", "Awake_Alert", "Deceased", "RACE", "tbi_severity",
    "Mechanism_Injury", "Marshall",
     "uni_incidence", "bi_incidence", "any_incidence" )
df$tbi_severity <- as.factor(df$tbi_severity)
df$RACE <- as.factor(df$RACE) %>%
relevel("White")
str(df)
```

# Incidence: Any, Unilateral, and Bilateral Pupil Abnormalities
## Print all Chi Square Values for Categorical Variables
```{r}
chisq.test(table(df$any_incidence, df$tbi_severity))$expected # < 5
chisq.test(table(df$poor_npi_incidence, df$tbi_severity))$expected
chisq.test(table(df$npi_diff_incidence, df$tbi_severity))$expected
chisq.test(table(df$size_diff_incidence, df$tbi_severity))$expected
chisq.test(table(df$uni_any_incidence, df$tbi_severity))$expected # < 5
chisq.test(table(df$bi_incidence, df$tbi_severity))$expected # < 5
```

## Print all Fisher Tests if Expected Count < 5 for Chi Square
```{r}
fisher.test(df$any_incidence, df$tbi_severity)
fisher.test(df$uni_any_incidence, df$tbi_severity)
fisher.test(df$bi_incidence, df$tbi_severity)
```

# Categorical Variables: Sex, Race, Mechanism Injury, Bolt, EVD, Cranial Surgery, Discharge Disposition, Unfavorable Outcome, Death, Awake_Alert
## Print all Chi Square Values for Categorical Variables
```{r}
chisq.test(table(df$SEX, df$tbi_severity))$expected
chisq.test(table(df$RACE, df$tbi_severity))$expected # < 5
chisq.test(table(df$Mechanism_Injury, df$tbi_severity))$expected # < 5
chisq.test(table(df$Bolt, df$tbi_severity))$expected # < 5
chisq.test(table(df$EVD, df$tbi_severity))$expected # < 5
chisq.test(table(df$Crani_Surgery, df$tbi_severity))$expected # < 5
chisq.test(table(df$Discharge_Disposition, df$tbi_severity))$expected # < 5
chisq.test(table(df$Unfavorable_Outcome, df$tbi_severity))$expected
chisq.test(table(df$Deceased, df$tbi_severity))$expected # < 5
chisq.test(table(df$Awake_Alert, df$tbi_severity))$expected # < 5
```
## Print all Fisher Tests if Expected Count < 5 for Chi Square (Race, Mechanism Injury, Bolt, EVD, Cranial Surgery, Discharge Disposition, Death, Awake_Alert)
```{r}
fisher.test(df$RACE, df$tbi_severity)
fisher.test(df$Mechanism_Injury, df$tbi_severity)
fisher.test(df$Bolt, df$tbi_severity)
fisher.test(df$EVD, df$tbi_severity)
fisher.test(df$Crani_Surgery, df$tbi_severity)
fisher.test(df$Discharge_Disposition, df$tbi_severity, simulate.p.value = TRUE)
fisher.test(df$Deceased, df$tbi_severity)
fisher.test(df$Awake_Alert, df$tbi_severity)
```
## Frequency Kruskal-Wallis Test
```{r}
kruskal.test(any_burden ~ tbi_severity, data = df)
kruskal.test(poor_npi_burden ~ tbi_severity, data = df)
kruskal.test(npi_diff_burden ~ tbi_severity, data = df)
kruskal.test(size_diff_burden ~ tbi_severity, data = df)
kruskal.test(uni_any_burden ~ tbi_severity, data = df)
```
# Wilcoxon-Mann-Whitney test
## Any Burden
```{r}
library(WMWssp)
UO_df<- filter(df, Unfavorable_Outcome == 1)
FO_df<- filter(df, Unfavorable_Outcome == 0)
x1 <- FO_df$any_burden
x2 <- UO_df$any_burden

WMWssp(x1, x2, alpha = 0.05, power = 0.988, t = 85/131)
```
# Wilcoxon-Mann-Whitney test
## Unilateral Burden
```{r}
x1 <- FO_df$uni_any_burden
x2 <- UO_df$uni_any_burden
WMWssp(x1, x2, alpha = 0.05, power = 0.67, t = 85/131)
```
# Wilcoxon-Mann-Whitney test
## Bilateral Burden
```{r}
x1 <- FO_df$bi_burden
x2 <- UO_df$bi_burden
WMWssp(x1, x2, alpha = 0.05, power = 0.9945, t = 85/131)
```


