---
title: "Univariate Analysis"
author: Divya Veerapaneni MS4, Ong Lab
input: Consolidated_Study_Cohort_Traits.xlsx - excel sheet with full TBI cohort details including demographic data, pupil metrics, incidence and frequency of abnormal pupil phenotypes 
output: 
  html_document:
    df_print: paged
---
#### Input file: logistic_regression_data_137_final.xlsx 
Excel sheet with preprocessed data for final 137 TBI patients including demographic data, pupil metrics, incidence and burden of abn pupil phenotypes 

```{r}
#import statements
library(readxl)
library(gtsummary)
library(broom)
library(lmtest)
library(tidyverse)
library(purrr)

#input dataframe
df <- read_excel("/Users/divs/Box/1-BMC Smartguards/10-Processing and Visualization/8-TBI Pupillometry/Data/Consolidated_Study_Cohort_Traits.xlsx")

#variables of interest in input file
factor_variables = list("Unfavorable_Outcome", "Awake_Alert", "Deceased", "RACE_Black", "tbi_severity",
    "Mechanism_Injury", "Marshall", "Rotterdam",
     'any_incidence', 'poor_npi_incidence', 'npi_diff_incidence',
       'size_diff_incidence', 'uni_any_incidence', 'stage1u_incidence',
       'stage2u_incidence', 'stage3u_incidence', 'stage4u_incidence',
       'bi_incidence', 'stage1b_incidence', 'stage2b_incidence',
       'stage3b_incidence' )
#dft$RACE <- relevel(df$RACE, ref="White")
df$RACE <- as.factor(df$RACE) %>%
relevel("White")

colnames(df)
```

# Effect of Possible Risk Factors on Any Frequency
Includes beta coefficient, 95%Cl, and p values
```{r}
explanatory_vars <- c("AGE", "SEX", "RACE_Black", "RACE", "tbi_severity", "GCS","Mechanism_Injury", "Bolt", "EVD", "Crani_Surgery", "Marshall", "Rotterdam", "Time_to_most_severe_CT", "time_to_first_obs", "lower_npi","average_npi", "npi_diff", "size_diff", "time_to_first_obs")

univ_tab <- df %>% 
  dplyr::select(explanatory_vars, any_burden) %>% ## select variables of interest

  tbl_uvregression(                         ## produce univariate table
    method = lm,                           ## define regression want to run (generalized linear model)
    y = any_burden,                            ## define outcome variable
  )
univ_tab
```

# Effect of Risk Factors on Unilateral Pupil Abnormalities
Includes beta coefficient, 95%Cl, and p values
```{r}

univ_tab <- df %>% 
  dplyr::select(explanatory_vars, uni_any_burden) %>% ## select variables of interest

  tbl_uvregression(                         ## produce univariate table
    method = lm,                           ## define regression want to run (generalized linear model)
    y = uni_any_burden,                            ## define outcome variable
  )
univ_tab
```

# Effect of Risk Factors on Bilateral Pupil Abnormalitie
Includes beta coefficient, 95%Cl, and p values
```{r}

univ_tab <- df %>% 
  dplyr::select(explanatory_vars, bi_burden) %>% ## select variables of interest

  tbl_uvregression(                         ## produce univariate table
    method = lm,                           ## define regression want to run (generalised linear model)
    y = bi_burden,                            ## define outcome variable
  )
univ_tab
```

# Stages Burden Univariate Analysis

```{r}
univ_tab <- df %>% 
  dplyr::select(explanatory_vars, stage1u_burden) %>% ## select variables of interest

  tbl_uvregression(                         ## produce univariate table
    method = lm,                           ## define regression want to run (generalised linear model)
    y = stage1u_burden,                            ## define outcome variable
  )
univ_tab
```

```{r}
univ_tab <- df %>% 
  dplyr::select(explanatory_vars, stage2u_burden) %>% ## select variables of interest

  tbl_uvregression(                         ## produce univariate table
    method = lm,                           ## define regression want to run (generalised linear model)
    y = stage2u_burden,                            ## define outcome variable
  )
univ_tab
```

```{r}
univ_tab <- df %>% 
  dplyr::select(explanatory_vars, stage3u_burden) %>% ## select variables of interest

  tbl_uvregression(                         ## produce univariate table
    method = lm,                           ## define regression want to run (generalised linear model)
    y = stage3u_burden,                            ## define outcome variable
  )
univ_tab
```

```{r}
univ_tab <- df %>% 
  dplyr::select(explanatory_vars, stage1b_burden) %>% ## select variables of interest

  tbl_uvregression(                         ## produce univariate table
    method = lm,                           ## define regression want to run (generalised linear model)
    y = stage1b_burden,                            ## define outcome variable
  )
univ_tab
```

```{r}
univ_tab <- df %>% 
  dplyr::select(explanatory_vars, stage2b_burden) %>% ## select variables of interest

  tbl_uvregression(                         ## produce univariate table
    method = lm,                           ## define regression want to run (generalised linear model)
    y = stage2b_burden,                            ## define outcome variable
  )
univ_tab
```

# Logistic Regression Univariate Analysis for Any Incidence vs Age, Sex, Race, and TBI Severity
Includes OR, 95%Cl, and p values
```{r}

models <- explanatory_vars %>%       # begin with variables of interest
  str_c("any_incidence ~ ", .) %>%  # combine each variable into formula
  
  # iterate through each univariate formula
  map(                               
    .f = ~glm(                       # pass the formulas one-by-one to glm()
      formula = as.formula(.x),      # within glm(), the string formula is .x
      family = "binomial",           # specify type of glm (logistic)
      data = df)) %>%          # dataset
  
  # tidy up each of the glm regression outputs from above
  map(
    .f = ~tidy(
      .x, 
      exponentiate = TRUE,           # exponentiate 
      conf.int = TRUE)) %>%          # return confidence intervals
  
  # collapse the list of regression outputs in to one data frame
  bind_rows() %>% 
  
  # round all numeric columns
  mutate(across(where(is.numeric), round, digits = 2))
models
```

